### 2014-10-22 thesis/method/classification.tex
Multi-layer perceptron and the conclusion of the chapter

\section{Multi-layer Perceptron (MLP)}

\section{Conclusion}

>>>>

\section{Multi-layer Perceptron (MLP)}
Multi-layer perceptrons (MLPs) are a generalised form of the
\textit{perceptron}, first
studied extensively by Rosenblatt \todo{add citation entry}. The original
motivation of the perceptron 
was to develop a very simple artificial neural network to model the
behaviour of neurons in the brain. The perceptron model for a two-class
problem is a linear combination of the features: that is,
$h(\mathbf{x}) = \sum w_i x_i$ where the $w_i$ are the weights of the model.
There is an input node, called a \textit{neuron}, for each feature, and the
weights are applied to the connections between these input neurons to the
output.
Usually, a constant term of 1 called the \textit{bias} is also added to the
model for an additional trainable input. To classify an item, we substitute it
into $h(\mathbf{x})$; if the value is above a particular threshold, we label it
as class 1, else it belongs in class 2. For convenience, we can let this
threshold be 0. Training the perceptron involves the following steps:
\begin{enumerate}
\item Start with random numbers for the weights $w_i$.
\item For each training example $\mathbf{x_i}$: calculate $h(\mathbf{x_i})$.
If it is correctly classified, do not update the weights.
If it is incorrectly classified, add a multiple of the misclassified
instance to the current weights using
$\mathbf{w} = \mathbf{w} + \lambda \mathbf{x_i}$, where $\lambda$ is a small
constant.
\item Repeat the previous step until all training examples are correctly
classified.
\end{enumerate}
If the data are linearly separable, the perceptron is guaranteed to find a
hyperplane that separates the two classes. However, it is not able to do this
when there is a non-linear decision boundary -- for that we will need the
multi-layer perceptron.

The \textit{multi-layer perceptron}, also known as a
\textit{feed-forward neural network}, is a type of artificial neural network
that generalises the perceptron model outlined above. Instead of only one layer
of inputs, MLPs consist of more than one layer, with the outputs from previous
layers serving as inputs to the next layer. At each layer, the inputs are
combined linearly and then transformed nonlinearly for input into the next
layer. For a network with one layer of transformations between the input and
output layers -- that is, one ``hidden'' layer -- the mathematical description
is:
\begin{equation*}
y = \sum_j w^{(2)}_j f_j \left(\sum_i w^{(1)}_i x_i\right)
\end{equation*}
The $w$ and the $f_j$ represent the linear combinations and the nonlinear
transformations respectively, and $y$ is the predicted outcome.
The nonlinearity of the transformations ensures
that we can classify data which are not linearly separable. Figure
\ref{fig:mlp} shows an MLP with one hidden layer.

\begin{figure}[h]
\label{fig:mlp}
\caption{}
\centering
%\includegraphics[scale=0.95]{images/a2-nn}
\end{figure}

MLPs can be trained in a number of ways, the most common of which is
\textit{backpropagation}. Suppose that we have the network above, with one
input, hidden and output layer respectively. From the perceptron learning
algorithm, we can update the weights of the hidden neurons leading to the
output neuron, but we do not know what the outputs of these hidden neurons are.
Backpropagation solves this problem by updating the weights of the various
connections to the hidden units based on the strength of the unit's
contribution to the final prediction, using a mathematical optimisation
algorithm called \textit{gradient descent}. This requires taking derivatives,
so the threshold function we used for the perceptron is usually replaced with
the sigmoid function $f(x) = \frac{1}{1+e^{-x}}$. We also define an error
function $E=\frac{1}{2}(\text{actual} - \text{predicted})^2$ -- the one most
commonly used -- which is
minimised in the gradient descent process. Updating the weights for each
connection in the network now involves finding the value of the derivative,
multiplying it by a small constant between 0 and 1 called the
\textit{learning rate}, and subtracting this value from the current weight
value, until the minimum for the error function is reached. Note that the
error function is quadratic, so it is guaranteed to have a minimum. The
updating of the weights is continued until the minimum is reached or if we
are ``close enough'' to the minimum.

It can be shown that MLPs with a single hidden layer are powerful enough to
model any continuous function, provided there are enough neurons in that
hidden layer \cite{Hand2001}. However, as they must iterate over the entire
data set many times, and due to the sometimes slow convergence of gradient
descent, MLPs are often slower to train than other learning algorithms we
have described, such as decision trees. We use them in this work on LOS
classification as we would like to see how they perform in modelling the
complex relationships between various medical attributes.

\section{Conclusion}
In this chapter we have outlined all of the learning algorithms that we will
be evaluating with respect to our classification problem, from the simple
0R and 1R to Na\"{i}ve Bayes, decision trees, logistic regression,
$k$-nearest neighbour (and extensions K* and Ranked Distance), to the more
sophisticated support vector machines and multi-layer perceptrons. The reason
for our inclusion of so many algorithms is two-fold. Firstly, many of
these algorithms have not been extensively applied in various medical settings
to the problem of LOS prediction. Using a wide range of algorithms from various
paradigms will enable us to see what algorithms perform better than others.
Secondly, evaluating several classifiers will allow us to make comparisons
between different feature selection methods that are independent of the
classifier and give us more insight into the nature of the data set and of the
problem.
